---
title: "Minor Allele Frequency"
author: "Jeffrey Long"
date: "11/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Minor Allele Frequency

Using the supplied file, create two outputs.


### Get and Read in Data

Getting Data:

```{bash eval=FALSE}
cp /Users/jeffreylong/Desktop/sampleFile\ 1.txt /Users/jeffreylong/r/minorAlleleFrequency/sampleFile\ 1.txt
```

Reading Data:

```{r}
library(tidyverse)
filename <- "~/r/minorAlleleFrequency/sampleFile\ 1.txt"
data <- read_tsv("sampleFile\ 1.txt")
```

### 1. Calculate minor allele frequency (MAF) for each SNP.

Where MAF = minimum(reference allele frequency, alternate allele frequency) 
Reference (or Alternate) allele frequency is defined by 
reference (or alternate) allele counts divided by the 
sum of reference and alternate allele counts for a given SNP. 
Code should print “SNPname” list and its corresponding MAF. 


```{r}
# Calculate and Output Solution for Code 1
result <- data %>%
  rowwise() %>%
  mutate(MAF = min(c(ReferenceAlleleCount,AlternateAlleleCount)) / (ReferenceAlleleCount+AlternateAlleleCount)) %>%
  select(SNPname,MAF)

result
```

### 2. Filter MAFs for range 0.02-0.3

Calculate the number of SNPs that has a minor allele 
frequency between 0.02 and 0.3. i.e., 0.02 <= MAF <= 0.3

Code should print out the file name and the number of SNPs 
matching the criterion.


```{r}
result_filtered <- result %>%
  filter(MAF >= 0.02 & MAF<= 0.3)

basename(filename)
nrow(result_filtered)
```


## Coding Environment
```{r}
sessionInfo()
```
